---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
library(GGally)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
load("movies.Rdata")
head(movies)
```

* * *

## Part 1: Data

### Context of the study

This project is based on a fictitious scenario - one where I've been hired as a data scientist at Paramount pictures. The data, sourced by Paramount, presents numerous variables on movies such as audience and critic ratings. Paramount endeavors to gather insights into determining the acclaim of a film and other novel patterns or ideas. The data set is comprised of 651 randomly sampled movies produced and released before 2016.

### Sampling Design

```{r}
movies$studio %>%
  unique() %>%
  head(10)

movies %>%
  summarise(max_year = max(thtr_rel_year), min_year = min(thtr_rel_year))
```

Some of the film studios in the dataset are Indomina Media Inc., Warner Bros. Pictures, and Columbia Pictures. The majority of these studios are based in the US, and since the date ranges between the years 2014 and 1970. Most of these film studios, including ones by Sony Corporation, headquartered in Tokyo, are shot in California ^[[Wikipedia - Film industry](https://en.wikipedia.org/wiki/Film_industry)]. Furthermore, US-based film studios were inexperienced in film productions abroad and often sold the rights to their films to foreign parties ^[[Wikipedia - Cinema of the United States](https://en.wikipedia.org/wiki/Cinema_of_the_United_States)]. It is thereby safe to assume such studios targeted a US-based audience. However, it should be noted that ratings produced by rotten tomatoes and IMDb tend to be based on a global audience.

### Scope of Inference

The data collected is observational and captures all the significant factors of a movie such as its genre, runtime, and whether or not the movie is in the Top 200 Box Office list on BoxOfficeMojo. Furthermore, The data is also a random sample of the movies. Since no strata, blocks, quotas, nor clusters are defined, a simple random sample can be assumed. Since this data does not result from an experimentation study, nor are there any variables that are blocked nor controlled, we cannot infer causation from the data.


### Generalizability

```{r}
year_n_movies <- movies %>%
  group_by(thtr_rel_year) %>%
  summarize(n_movies = n())

ggplot(data = year_n_movies, aes(x = thtr_rel_year, y = n_movies)) +
  geom_bar(stat='identity') +
  ggtitle("Number of movies released in theatre per year (Paramount Dataset)")
``` 

As according to the bar graph above, over 50% of the years have below 20 movies. Therefore, we cannot generalize these results to all U.S. movies released in each year. However, we can generalize this to a random sample of all U.S. movies as the dataset consists of 651. This is well below 10% of all movies released in the U.S. and it is a random sample, so we can safely assume each movie is independent of one another. 

* * *

## Part 2: Research question

Since Paramount is likely to intend to capitalize on popular movies, it leads us to the following research question:

> What are the driving factors for determining the popularity of a movie?

A popular movie implies higher box office sales resulting in more revenue for the company. Paramount is likely to benefit in the form of ticket sales, DVD sales, and licensing the movie to third party vendors. Though it would be ideal to capture the popularity of a movie in terms of its financial gain, due to the nature of this dataset, we cannot do so.

Such a prediction would also allow Paramount to decide which movie ideas to invest in based on its `genre`, `runtime`, `mpaa_rating` and other such factors available prior to the release of a movie. Furthermore, this project will be using the `imdb_rating` as our response variable. The IMDb is a Amazon owned online movie database that provides information related to films and television programs. As of January 2020, 83 million registered users ^[[Press Room - IMDb](https://www.imdb.com/pressroom/stats/)]. Since customers are likely to refer to IMDb ratings as a credible source for movie recommendation, a high rating is important.

The IMDb originally used the following formula to calculate their weighted rating ^[[Wikipedia - IMDb](https://en.wikipedia.org/wiki/IMDb)]:

\[
  W = \frac{R \times v + C \times m}{v + m}
\]

Where,

* $W$ = weighted rating
* $R$ = average for the movie as a number from 1 to 10 (mean) = (Rating)
* $v$ = number of votes for the movie = (votes)
* $m$ = minimum votes required to be listed in the Top 250 (currently 25,000)
* $C$ = the mean vote across the whole report (currently 7.0)

This means that an influential driver for higher ratings is a large number of top votes (preferably above 8). Furthermore, since this project is primarily concerned with significant variables, the p-value will be considered.

* * *

## Part 3: Exploratory data analysis

The explanatory variables of interest in our analysis are as follows:

| title            | Title of movie                                                                                                                                                                            |
|------------------|-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------|
| `genre`            | Genre of movie (Action & Adventure, Comedy, Documentary, Drama, Horror, Mystery & Suspense, Other)                                                                                        |
| `runtime`          | Runtime of movie (in minutes)                                                                                                                                                             |
| `mpaa_rating`      | MPAA rating of the movie (G, PG, PG-13, R, Unrated)                                                                                                                                       |
| `studio``           | Studio that produced the movie                                                                                                                                                            |
| `thtr_rel_year`    | Year the movie is released in theaters                                                                                                                                                    |
| `thtr_rel_month`   | Month the movie is released in theaters                                                                                                                                                   |
| `thtr_rel_day`     | Day of the month the movie is released in theaters                                                                                                                                        |
| `dvd_rel_year`     | Year the movie is released on DVD                                                                                                                                                         |
| `dvd_rel_month`    | Month the movie is released on DVD                                                                                                                                                        |
| `dvd_rel_day`      | Day of the month the movie is released on DVD                                                                                                                                             |
| `best_pic_nom`     | Whether or not the movie was nominated for a best picture Oscar (no, yes)                                                                                                                 |
| `best_pic_win`     | Whether or not the movie won a best picture Oscar (no, yes)                                                                                                                               |
| `best_actor_win`   | Whether or not one of the main actors in the movie ever won an Oscar (no, yes) – note that this is not necessarily whether the actor won an Oscar for their role in the given movie       |
| `best_actress_win` | Whether or not one of the main actresses in the movie ever won an Oscar (no, yes) – not that this is not necessarily whether the actresses won an Oscar for their role in the given movie |
| `best_dir_win`     | Whether or not the director of the movie ever won an Oscar (no, yes) – not that this is not necessarily whether the director won an Oscar for the given movie                             |


This analysis will not consider the following as explanatory variables. These correspond to so some form of ratings or ranking on Imdb, rotten tomatoes or BoxOfficeMojo:

* `imdb_num_votes`: Number of votes on IMDB
* `critics_rating`: Categorical variable for critics rating on Rotten Tomatoes (Certified Fresh, Fresh, Rotten)
* `critics_score`: Critics score on Rotten Tomatoes
* `audience_rating`: Categorical variable for audience rating on Rotten Tomatoes (Spilled, Upright)
* `audience_score`: Audience score on Rotten Tomatoes
* `top200_box`: Whether or not the movie is in the Top 200 Box Office list on BoxOfficeMojo (no, yes) 

Based on domain knowledge, ratings on other websites wouldn't be available prior to the release of the movie and such variables are likely to have high collinearity with the response variable which will result in incorrect estimators for other factors in this analysis. We can use the a pairplot to verify our assumption:

```{r}
# numerical variables
ggpairs(movies, columns = c('imdb_rating','imdb_num_votes','critics_score','audience_score'))
```

As mentioned earlier, some numerical rating variables such as `audience_score` and `critics_score` on rotten tomatoes tend to result in a high correlation coefficient. Thus, they have a strong linear association with the response variable, which could result in biased estimators for other variables. Furthermore, these variables won't be available to us prior to the release of a movie.

```{r}
# categorical variables
ggpairs(movies, columns = c('imdb_rating', 'critics_rating', 'audience_rating', 'top200_box'))
```

In terms of categorical variables, we can use anova to verify if the variability 

* * *

## Part 4: Modeling



* * *

## Part 5: Prediction



* * *

## Part 6: Conclusion

## References
