---
title: "Modeling and prediction for movies"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `movies`. Delete this note when before you submit 
your work. 

```{r load-data}
load("movies.Rdata")
head(movies)
```



* * *

## Part 1: Data

### Context of the study

This project is based on a fictitious scenario - one where I've been hired as a data scientist at Paramount pictures. The data, sourced by Paramount, presents numerous variables on movies such as audience and critic ratings. Paramount endeavors to gather insights into determining the acclaim of a film and other novel patterns or ideas. The data set is comprised of 651 randomly sampled movies produced and released before 2016.

### Sampling Design

```{r}
movies$studio %>%
  unique() %>%
  head(10)

movies %>%
  summarise(max_year = max(thtr_rel_year), min_year = min(thtr_rel_year))

```

Some of the film studios in the dataset are Indomina Media Inc., Warner Bros. Pictures, and Columbia Pictures. The majority of these studios are based in the US, and since the date ranges between the years 2014 and 1970. Most of these film studios, including ones by Sony Corporation, headquartered in Tokyo, are shot in California ^[[Wikipedia - Film industry](https://en.wikipedia.org/wiki/Film_industry)]. Furthermore, US-based film studios were inexperienced in film productions abroad and often sold the rights to their films to foreign parties ^[[Wikipedia - Cinema of the United States](https://en.wikipedia.org/wiki/Cinema_of_the_United_States)]. It is thereby safe to assume such studios targeted a US-based audience. However, it should be noted that ratings produced by rotten tomatoes and IMDb tend to be based on a global audience.

### Scope of Inference

The data collected is observational and captures all the significant factors of a movie such as its genre, runtime, and whether or not the movie is in the Top 200 Box Office list on BoxOfficeMojo. Furthermore, The data is also a random sample of the movies. Since no strata, blocks, quotas, nor clusters are defined, a simple random sample can be assumed. Since this data does not result from an experimentation study, nor are there any variables that are blocked nor controlled, we cannot infer causation from the data.

We can, therefore, generalize these results to all U.S. movies released between 1970 and 2014.


* * *

## Part 2: Research question

Since Paramount is likely to intend to capitalize on popular movies, it leads us to the following research question:

> What are the driving factors for determining the popularity of a movie?

A popular movie implies higher box office sales resulting in more revenue for the company. Paramount is likely to benefit in the form of ticket sales, DVD sales, and licensing the movie to third party vendors. Though it would be ideal to capture the popularity of a movie in terms of its financial gain, due to the nature of this dataset, we cannot do so.

This project will be using the `imdb_rating` as our response variable. The IMDb is a Amazon owned online movie database that provides information related to films and television programs. As of January 2020, 83 million registered users ^[[Press Room - IMDb](https://www.imdb.com/pressroom/stats/)]. Since customers are likely to refer to IMDb ratings as a credible source for movie recommendation, a high rating is important.

The IMDb originally used the following formula to calculate their weighted rating ^[[Wikipedia - IMDb](https://en.wikipedia.org/wiki/IMDb)]:

\[
  W = \frac{R \times v + C \times m}{v + m}
\]

Where,

* $W$ = weighted rating
* $R$ = average for the movie as a number from 1 to 10 (mean) = (Rating)
* $v$ = number of votes for the movie = (votes)
* $m$ = minimum votes required to be listed in the Top 250 (currently 25,000)
* $C$ = the mean vote across the whole report (currently 7.0)

This means a major driver for higher rating are a large number of high votes (preferably above 8). Furthermore, since this proejct is primarily sigificant variables, the p-value will be considered.

* * *

## Part 3: Exploratory data analysis



* * *

## Part 4: Modeling



* * *

## Part 5: Prediction



* * *

## Part 6: Conclusion

## References
